title

### 2024-12-18 index.qmd
long time no commit

# code to recode levels "-99" in factor vars
data <- data |> 
  mutate(across(where(is.factor), ~forcats::fct_recode(., NULL = "-99")))

@@

@@

Election administration officials make efforts to sustain public trust and confidence in the fairness and accuracy of elections, and attempt to boost such confidence where ever it may be deprived. Concerns for safety have developed among election staff and voters in more recent elections. Regular measures are taken to enhance the *trustworthiness*[^1] of the electoral process through practices meant to improve the conduct, transparency, or overall administration of elections in the United States.

@@

In this paper, I report results from a recent survey experiment administered to test whether publicized efforts to recruit veterans to work as election staff and volunteers would ease election anxieties. Results of the survey experiment support the hypothesis that emphasizing veterans as the target of election worker recruitment efforts eases pre-elections anxieties to an extent. In particular, among those in the treatment condition, confidence that the elections would be fair and accurate was higher in comparison to those in the control. In addition, expectations of electoral fraud were lower, as were concerns for voter safety compared to those who read a control vignette. Notable is that there was a significant difference in confidence, and lower expectations of fraud, among those in the treatment condition who believe that the 2020 election was illegitimate.

@@

Suffice to say, pre-election anxiety consists of more than confidence in fairness and accuracy of vote counts in light of added safety concerns. It is not hard to recognize that increased tension in the pre-election period makes for a volatile political environment.

@@

Assessing the public's *trust* in elections has not been straightforward. Inquiry into public trust in elections has been approached by scholars of political science in many different ways [@cook2005], often distinguishable by the scope of the research question and more or less constrained by the particular conception of public trust. Quite often, trust in election administration is conflated with trust in government writ large, government legitimacy, government or system responsiveness, or even satisfaction with democracy [@daniller2019]. At this level, not only is the level of public trust in elections sometimes vague, but there's little consideration over the difference between such attitudes pre-election and post-election. In contrast, a considerable amount of research tends to conceive of public trust in accordance with the institution in question [@atkeson2007; @hooghe2018].

Election administration is just one part of the larger set of institutions which form the electoral system. As such, the performance of the institution along with the rest "...lends credibility to the outcome of an election: whether it is considered by citizens and the international community to be fair and legitimate." [@stewart2022, 236].

Trust and confidence in the conduct of elections concerns aspects of elections that fall squarely within the institution of election administration. At this level, for instance, public trust is ascertained by capturing assessments about the perceived accuracy of vote counts (e.g., whether votes are/were counted as intended). Or in other words, public trust is ascertained by the extent the public is confident in the administration of elections.

Intuitively, enhancing public trust in elections would best be accomplished by enhancing the *trustworthiness* of the institution, i.e., consistently doing the things that election officials already regularly do come election time. However, trust and confidence in elections has become ever more precarious over the last few election cycles. Especially considering public polling data since 2000 shows that confidence that votes were, or would be, counted as intended was in a consistent decline despite efforts towards bolstering election integrity and trustworthiness [@sances2015]. This is even more pronounced considering the role that partisanship has had on such confidence over accuracy of vote count [@sances2015; @stewart2022].

@@

As such, it is questionable whether we can characterize public trust in the pre-election period as the same trust after the election results have come out[^3]. The former is *anticipatory*---i.e., the kind that is more or less anxious given the uncertainties surrounding the election. The latter is *evaluative*---a judgement discerned empirically in hindsight after the election event has occurred. The evaluative form of trust lends to inquiries interested in public opinion on regime legitimacy, satisfaction with democracy, and perceived responsiveness [@daniller2019]. This study focuses primarily on that *anticipatory* kind of trust and confidence, which speaks more to electoral anxieties over the institution of election administration. Not to mention that by definition, one feels anxiety over impending uncertainties located in the future.

The primary interest of this study is public trust in election administration prior to the election, but differentiates between confidence and concern, or otherwise trust v. anxiety. The two are inversely related, at least conceptually. As such, we might also expect an inverse relationship between reported confidence and reported concern over the different aspects of electoral administration and process. Though that leads to the conceptual wrinkle that distinguishes between trust, distrust, and anxiety, which is addressed later.

*election anxieties*[^4], which is distinguished into 
two categories: fairness and accuracy, and voter safety. Of course, the direction of electoral outcomes are a particular source of electoral anxiety, but only the election outcome itself can resolve any nail-biting stress over its anticipated direction. In contrast, the anxieties of interest here are able to be relaxed to some extent prior to the election given some intervention.

Interventions aimed at improving public trust, or rather, aimed at easing election anxieties may not be within the usual repertoire election officials draw upon. Regardless of the measures taken by election officials to boost public confidence in the *trustworthiness* of election administration (e.g., conducting audits, testing election machines), public *trust* and confidence in elections more generally is apt to shift dramatically post-election based on factors such as partisanship, elite rhetoric, particular state policies, and more [@nadeau1993; @carter2024; @coll2024a]. Moreover, prior research has found that evaluation of election workers themselves are an important factor when it comes to levels of public confidence in the electoral process [@hall2007; @hall2009; @claassen2008]. Such studies focused on the quality of the voter experience with reference to the interaction between voter and election worker.

Beyond the general competence of election workers, however, the quality of the voter experience may be influenced merely by impressions about *who* comprises election staff and volunteers. Political and other social science researchers have recognized for some time the power that group identity can have over attitudes and perception [@xiao2016; @xiao2012; @vanbavel2021]. As such, we can expect that information identifying the particular groups being recruited to serve as election staff will be enough to improve trust and confidence in election administration, i.e., ease election anxieties. It was hypothesized that information on recruitment efforts targeting military veterans to work as election staff would ease election anxieties. The next section elaborates further on why military veterans are of particular interest.

@@

The reason why veterans arose as a special subset of the population to consider for this inquiry is because there was a sudden push to target veterans for recruitment efforts that arose shortly after the events on Capitol Hill on January 6th, 2021. After the 2020 election, large efforts were made to recruit military veterans and their families to work or volunteer as election staff [@wetheveterans2022; @nflfootballoperations2022]. Prior to that point, young people were sometimes given special mention as targets of election worker recruitment efforts[^5] [@powerthepolls2020; @herndon2020]. Generally, however, recruitment efforts cast a wide net, indiscriminate of who applies [@ross2020; @conde2020].

One can speculate that the motivation to associate military veterans with civic engagement and democracy may be intended to counter negative perceptions and impressions given by the proportion of veteran service members arrested for taking part in the events on January 6th [@milton2021; @loewenson2023; @jensen2022a]. Especially with regard to research demonstrating that willingness to support violent efforts to overturn election results[^6] (in support of Trump) is, on average, more common among veterans than among matched samples of non-veterans [@pape2024]. This is in addition to a strengthened association portrayed in media outlets between military veterans and militias [@steinhauer2020]. Although such media output reeks of the contemporary political context, prior research has substantiated such a connection between veterans and militia groups. @cooter2024 notes from her 3-year ethnographic fieldwork among Michigan militia members that, "...approximately 40% of militia leaders and 30% of members had previous military experience. Most of these veterans actively sought out such groups, as opposed to being recruited by them" [-@cooter2024; see also @cooter2013]. Thus, countering such associations by promoting a different image of veterans to the mass public and veterans alike seems like a reasonable motivation. Yet such speculation is just that.

@@

## Theory

## Research Design and Method

To test the theory that publicized efforts to recruit veterans to work as election staff and volunteers would ease election anxieties, a recent survey experiment was embedded in a survey developed and conducted by the Center for Democracy and Civic Engagement (CDCE) at the University of Maryland (UMD). The survey was fielded from August 29th, 2024 to September 18th, 2024 on a non-probability sample of 1,287 U.S. citizens 18 years of age or older.

The primary independent variable is the prospect of military veterans as poll workers. It was considered that veterans would elicit high degrees of approval and support for whatever cause or issue presented. For example, I'd expect high approval and support for recruiting veterans to baseball teams just as much as for recruiting them to work as election staff. Asking participants whether they supported programs intent on recruiting veterans to work as election staff would garner support regardless of the target group. Considering that support and admiration for veterans is generally high among the population [@kleykamp2023], then discerning the impact on one's confidence would require a survey experiment and questions designed to determine whether veterans have any special effect on electoral confidence and voter safety distinct from the generally high admiration observed most often.

Survey participants were randomly assigned to read one of two fabricated news articles (i.e., vignettes) about recruitment efforts for election jobs. Participants read either a treatment or control vignette, which was a fabricated news article about efforts in Maricopa County, AZ to recruit election staff and volunteers for the 2024 general election. The treatment vignette referred to a program designed to recruit veterans and their family members and describes an interviewee "Jordan Braxton" as an Army veteran; whereas the control vignette simply omitted any mention of veterans and their family members, and didn't describe "Jordan Braxton" as an Army veteran. Beyond those small differences and the headlines, the article vignettes are identical. Any effects can be attributed to the information about veterans in the treatment vignette[^7].

It should be noted that Maricopa County, AZ was chosen as the setting of the story in the vignettes due to the increased scrutiny levied toward election administration there after the 2020 election [@giles2021; @maricopacountyelectionsdepartment2022]. In addition, the survey includes a series of items which ask specifically about Maricopa County, AZ and a comparable series of questions about one's local area[^8].

Over the past two decades it became commonplace for national polls to gauge public confidence in election administration by asking some variety of the question, "How confident are you that your vote \[will be/was\] counted as you intended in the most recent election?" [@hall2009; @sances2015; @stewart2022]. In addition, since 2008, the Survey of the Performance of American Elections (SPAE) has included a good number of relevant questions to more thoroughly assess trust and confidence in election administration. Such questions inquire into the voter experience with the institution more directly. This study borrows, modifies, or takes inspiration from certain questions found within the 2022 SPAE and other survey items used by the Pew Research Center [@dunn2018]. These survey items comprise the dependent variables[^9].

The outcomes of interest (i.e., dependent variables) are broadly referred to as *election anxieties*, but more specifically refer to confidence over fairness and accuracy of the election administration; expectations of fraudulent activity intent on manipulating the electoral outcomes; and concern for voter safety, i.e., comfort voting in-person as it relates to expectations of violence, threats, or intimidation. The following presents multiple one-sided directional hypotheses. Previous research on public perception and regard for U.S. military veterans gives credence to the directional hypotheses [@kleykamp2018; @kleykamp2023].

The first outcome of interest regards *trust and confidence* in election administration, which includes the integrity and security of the electoral process, administration, technology, and people or organizations to fairly and accurately conduct elections. Survey items measured the degree of confidence that votes will be counted as voters intend, confidence that electoral systems are secure from technological threats, perceived commitment of election staff, confidence that outcomes will be fair, as well as confidence that the voting process will be fair.

The first hypothesis is as follows:

$H_{1}:$ Trust and confidence in election administration will be higher among those who were presented information about efforts to recruit veterans as election workers compared to those who were not

In total, six survey items measure different aspects of trust and confidence in election administration. These items were analyzed individually but were also transformed into a composite scale of trust and confidence.

The literature concerning public regard for veterans gives substantial reason to expect that publicized efforts to recruit veterans to work as election staff and volunteers would have the effect of easing election anxieties. As such, survey participants also responded to a series of questions that inquired into expectations of electoral fraud---i.e., fraudulent activity intent on manipulating the electoral outcome.

Note, however, that an *expectation* of electoral fraud implies a degree of distrust more so than an agitated state of anxiety over the *potential* for electoral fraud. Conceptually, this may be considered a particular kind of 'anxiety' given that is it pre-election (hence, an expectation), however, not necessarily. A person may be rather confident with regard to how much they distrust the integrity of electoral administration, resulting in such expectations.

For instance, the first item of the 5-item series asks participants to report their expectation that "There will be voter fraud, that is, people who are not eligible to vote will vote or vote more than once" on a 4-option response scale ranging from "Not likely at all" to "very likely". This statement, as well as the rest in the series, asserts an expected compromise of election integrity, which speaks to the trustworthiness of the institution. What the individual expects, more or less, is that electoral fraud will occur, implying that fraudulent attempts will be successful.

It was expected that the experimental stimulus would result in lower expectation for electoral fraud among those who read the treatment vignette compared to those who read the control vignette. The second hypothesis is,

$H_{2}:$ Expectations of electoral fraud will be lower among those who were presented information about efforts to recruit veterans as election workers compared to those who were not

Another outcome of interest regards concerns about voting in-person at a polling site, i.e., voter safety. That is, the extent people feel that voters are safe enough to cast a vote at polling sites free from violence, threats of violence, or intimidation. Two survey items use slightly different questions to inquire about the same essential issue—i.e., perceptions of physical safety at the polls. The third hypothesis follows,

$H_{3}:$ Concerns for voter safety be lower among those who are presented information about efforts to recruit veterans as election workers compared to those who are not

Another outcome of interest concerned how particular actions of election officials, or circumstances at polling sites, would impact the confidence of election integrity as well as voter safety *from the perspective of survey respondents*. Survey participants responded to a series of six statements, all prefaced with the same question, "Regardless of whether any of these are actually the case, how would the following impact your confidence in the fairness and accuracy of elections conducted this November?" The six statements were,

1.  Election officials test every machine used in the election to ensure they are secure.
2.  Election officials conduct audits of ballots after every election to confirm the results were accurate.
3.  Poll watchers affiliated with the political parties or candidates observe the election.
4.  Election staff and volunteers include military veterans and their family members from the community.
5.  Election staff and volunteers include lawyers from the community.
6.  Election staff and volunteers include college students from the community.

For each statement, survey participants responded by selecting one of five response options:

1.  Decrease confidence a lot
2.  Decrease confidence somewhat
3.  No impact on confidence
4.  Increase confidence somewhat
5.  Increase confidence a lot

It was expected that respondents would estimate stronger impact on their confidence given they were assigned to read the treatment vignette compared to those who were not. Thus, the hypothesis is,

$H_{4}:$ Respondents who read the treatment vignette will report stronger impact on their personal confidence in fairness and accuracy of election administration when veterans are stated to be included among election staff and volunteers compared to those who read the control vignette.

The extent people believe veterans are more trustworthy as election workers than non-veterans in general is difficult without comparable groups to consider. Recruitment efforts in 2022 targeted young students, lawyers, and military veterans [@wang2022]. Thus, lawyers and college students were selected as two groups for which responses could be compared between different statement versions.

The goal was to determine whether there is anything special in having veterans volunteer as election staff (i.e., poll workers) compared to other groups. That is, does it make any difference in terms of concerns over election security, fairness, and voter safety to include veterans as election staff and volunteers compared to other groups? The fourth hypothesis, irrespective of the experimental stimulus, reads,

$H_{5}:$ Respondents who read the treatment vignette will report stronger impact on their personal confidence in fairness and accuracy of election administration when veterans are stated to be included among election staff and volunteers compared to identical statements about lawyers or college students.

@@

@@

@@

[^2]: It is important to point out here that election administration is just one institution that interacts with many other institutions and processes which form the electoral system *writ large* [@stewart2022]. The way in which I discuss the institution of election administration, as well as the way in which I discuss and distinguish public trust in elections from broader conceptions of public trust, is borrowed substantively from @stewart2022's conceptual framework of the same. Distinctions between *trust* and *trustworthiness* and the notion of election administration as an institution are already apparent inspirations convenient for situating the theory of this research inquiry.

[^3]: There's little that could justify a conceptual distinction between public trust pre-election and public trust post-election. The temporal element renders the difference between pre- and post-election operative rather than conceptual.

>>>>

# code to recode levels "-99" in factor vars
data <- data |> 
  mutate(across(where(is.factor), ~forcats::fct_recode(., NULL = "-99"))) |> 
  # Reverse response options to show increase in *less concern*
  mutate(q25 = forcats::fct_rev(q25))

@@

# re-assign var labels to variables
# To quickly assign the variable labels, first create a named vector via
# deframe() with values as the variable labels and names as the variable names.
data_labels <- data_dict |>
  select(var, label_var) |> 
  deframe()

# Now assign the labels using the splice operator. Using the splice operator,
# labels are assigned via matching against the variable name, which means that
# variable order does not matter.
data <- data |> 
  labelled::set_variable_labels(!!!data_labels)

@@

Election administration officials make efforts to sustain public trust and confidence in the fairness and accuracy of elections, and attempt to boost such confidence where it may be deprived. Concerns for safety have developed among election staff and voters in more recent elections. Regular measures are taken to enhance the *trustworthiness*[^1] of the electoral process through practices meant to improve the conduct, transparency, or overall administration of elections in the United States.

@@

In this paper, I report results from a recent survey experiment administered to test whether publicized efforts to recruit veterans to work as election staff and volunteers would improve public trust in elections and ease election anxieties. Results of the survey experiment support the hypothesis that emphasizing veterans as the target of election worker recruitment efforts eases pre-elections anxieties to an extent. In particular, among those in the treatment condition, confidence that the elections would be fair and accurate was higher in comparison to those in the control. In addition, expectations of electoral fraud were lower, as were concerns for voter safety compared to those who read a control vignette. Notable is that there was a significant difference in confidence, and lower expectations of fraud, among those in the treatment condition who believe that the 2020 election was illegitimate.

[Insert why this is an important contribution]

This paper is structured as follows. First, I provide a brief background on public trust in election administration. Second, I synthesize a review of relevant literature with a focus on how political and social science has conceptualized and ascertained public trust in elections. Next, I supply reasoning for why military veterans are singled out as the relevant subset of the population in this study.

@@

Suffice to say, pre-election anxiety consists of more than confidence in fairness and accuracy of vote counts in light of added safety concerns. It is not hard to recognize that increased tension in the pre-election period makes for a volatile political environment. Sustaining trustworthiness in election administration is only more difficult in an environment where turnover of election workers increases and the struggle to recruit volunteers worsens in light of safety concerns. Since election worker performance is significant to public evaluation of elections, added safety concerns that drive out election staff and repel volunteers can only detract from trustworthiness of the institution.

@@

Trust and confidence in the conduct of elections concerns aspects of elections that fall squarely within the institution of election administration. At this level, for instance, public trust is ascertained by capturing assessments about the perceived accuracy of vote counts (e.g., whether votes are/were counted as intended). Or in other words, by the extent the public is confident in the accurate administration of elections.

Assessing the public's *trust* in elections has not been straightforward, however. Inquiry into public trust in elections has been approached by scholars of political science in many different ways [@cook2005; @hardin2004], often distinguishable by the scope of the research question and more or less constrained by the particular conception of public trust. Quite often, trust in election administration is conflated with trust in government *writ* large, government legitimacy, government or system responsiveness, or even satisfaction with democracy [@daniller2019]. At this level, not only is the level of public trust in elections sometimes vague, but there's little consideration over the difference between such attitudes pre-election and post-election. In contrast, a considerable amount of research tends to conceive of public trust in accordance with the institution in question [@atkeson2007; @hooghe2018].

Election administration is just one institution a part of the larger set of institutions which form the electoral system. As such, the performance of the institution along with the rest "...lends credibility to the outcome of an election: whether it is considered by citizens and the international community to be fair and legitimate." [@stewart2022, 236]. Intuitively, enhancing public trust in elections would best be accomplished by enhancing the *trustworthiness* of the institution, i.e., consistently doing the things that election officials already regularly do come election time. However, trust and confidence in elections has become ever more precarious over the last few election cycles. Especially considering public polling data since 2000 shows that confidence that votes were, or would be, counted as intended was in a consistent decline despite efforts towards bolstering election integrity and trustworthiness [@sances2015]. This is even more pronounced considering the role that partisanship has had on such confidence over accuracy of vote count [@sances2015; @stewart2022].

@@

As such, it is questionable whether we can characterize public trust in the pre-election period as the same trust after the election results have come out. The former is *anticipatory*---i.e., the kind that is more or less anxious given the uncertainties surrounding the election. The latter is *empirical*---a judgement discerned in hindsight after the experience of the election event has occurred. Evaluations of one's own trust and confidence in election administration based on the experience of voting is influenced by that very experience. As discussed by @stewart2022, not only was confidence in vote count accuracy influenced by the voter experience, but so too were evaluations of election officials [-@stewart2022, 242-243]. Not to mention the influence that the election results would also have on such evaluations. In other words, evaluations of trust and confidence post-election are influenced by one's interpretation of their experience. The subtle difference is simply the degree of uncertainty one feels in anticipation of the next election event. This study focuses primarily on that *anticipatory* kind of trust and confidence, which speaks more to those anxieties[^3] based on perceptions of the institution's trustworthiness than upon the particular voter experience.

Regardless of the measures taken by election officials to boost public confidence in the *trustworthiness* of election administration (e.g., conducting audits, testing election machines), public *trust* and confidence in elections more generally is apt to shift dramatically post-election based on factors such as partisanship, elite rhetoric, particular state policies, and more [@nadeau1993; @carter2024; @coll2024a]. Moreover, prior research has found that evaluation of election workers themselves are an important factor when it comes to levels of public confidence in the electoral process [@hall2007; @hall2009; @claassen2008]. Such studies focused on the quality of the voter experience with reference to the interaction between voter and election worker.

Beyond the general competence of election workers, however, the quality of the voter experience may be influenced merely by *a priori* impressions about *who* comprises election staff and volunteers. Political and other social science researchers have recognized for some time the power that group identity can have over attitudes and perception [@xiao2016; @xiao2012; @vanbavel2021]. As such, we can expect that information identifying the particular groups being recruited to serve as election staff will be enough to improve trust and confidence in election administration, and lessen expectations of electoral fraud. The next section elaborates on why military veterans are of particular interest in this regard.

@@

The reason why veterans arose as a special subset of the population to consider for this inquiry is because there was a sudden push to target veterans for recruitment efforts that arose shortly after the events on Capitol Hill on January 6th, 2021. After the 2020 election, large efforts were made to recruit military veterans and their families to work or volunteer as election staff [@wetheveterans2022; @nflfootballoperations2022]. Prior to that point, young people were sometimes given special mention as targets of election worker recruitment efforts [@powerthepolls2020; @herndon2020]. Generally, however, recruitment efforts cast a wide net, indiscriminate of who applies [@ross2020; @conde2020].

One can speculate that the motivation to associate military veterans with civic engagement and democracy may be intended to counter negative perceptions and impressions given by the proportion of veteran service members arrested for taking part in the events on January 6th [@milton2021; @loewenson2023; @jensen2022a]. Especially with regard to research demonstrating that willingness to support violent efforts to overturn election results (in support of Trump) is, on average, more common among veterans than among matched samples of non-veterans [@pape2024]. This is in addition to a strengthened association portrayed in media outlets between military veterans and militias [@steinhauer2020]. Prior research has substantiated such a connection between veterans and militia groups. @cooter2024 notes from her 3-year ethnographic fieldwork among Michigan militia members that, "...approximately 40% of militia leaders and 30% of members had previous military experience. Most of these veterans actively sought out such groups, as opposed to being recruited by them" [-@cooter2024; see also @cooter2013]. Thus, countering such associations by promoting a different image of veterans to the mass public and veterans alike seems like a reasonable motivation. Yet such speculation is just that.

@@

## Design and Measures

To test the theory that publicized efforts to recruit veterans to work as election staff and volunteers would improve trust in elections and ease election anxieties, a recent survey experiment was embedded in a survey developed and conducted by the Center for Democracy and Civic Engagement (CDCE) at the University of Maryland. The survey was fielded from August 29th, 2024 to September 18th, 2024 on a non-probability sample of 1,287 U.S. citizens 18 years of age or older. Respondents were randomly split into either the treatment (n = 650, 50.51%) control (n = 637, 49.49%) conditions[^Demographic breakdown of the sample are included in Appendix B.].

Survey participants read either a treatment or control vignette, which was a fabricated news article about efforts in Maricopa County, AZ to recruit election staff and volunteers for the 2024 general election. The treatment vignette referred to a program designed to recruit veterans and their family members and describes an interviewee "Jordan Braxton" as an Army veteran. The control vignette simply omitted any mention of veterans and their family members, and didn't describe "Jordan Braxton" as an Army veteran. Beyond those small differences and the headlines, the article vignettes are identical. Any effects can be attributed to the information about veterans in the treatment vignette[^5].

It should be noted that Maricopa County, AZ was chosen as the setting of the story in the vignettes due to the increased scrutiny levied toward election administration there after the 2020 election [@giles2021; @maricopacountyelectionsdepartment2022]. A block of survey items asked specifically about Maricopa County, AZ, followed by an identical block of items that asked the same questions about one's local area. Results between the items where the setting differs are compared and discussed.

### Trust in Elections Admin

Over the past two decades it became commonplace for national polls to gauge public confidence in election administration (i.e., voter confidence) by asking some variety of the question, "How confident are you that your vote \[will be/was\] counted as you intended in the most recent election?" [@hall2009; @sances2015; @stewart2022]. In addition, since 2008, the Survey of the Performance of American Elections (SPAE) has included a good number of relevant questions to more thoroughly assess trust and confidence in election administration. Such questions inquire into the voter experience with the institution more directly. This study borrows, modifies, or takes inspiration from certain question items found within the 2022 SPAE and other survey items from the Pew Research Center's 2018 American Trends Panel wave 38 [@dunn2018].

Note that, except for the item assessing voter confidence (i.e., votes counted as voters intend), most other question items from the 2022 SPAE inquire into the voter experience after the election has occurred. The items in this study, however, all inquire into confidence over one's expectations in anticipation of upcoming elections. In addition, other survey questionnaires ask similar questions that specifically contrast between a respondent's own local community, their state, and nationwide. This is in addition, or complementary to, items that distinguish between local area officials, state election officials, state government, and the U.S. federal government. The survey of this study, instead, focused on attitudes regarding election workers such as election officials, staff, and volunteers. As such, the items are utilized to comprise a component[^7] I broadly refer to as *trust in elections*.

[^7]: The methodological concerns between the use of (unit weighted) sum scores compared to the use of estimated factor scores for Likert scales composed of multiple Likert items are legion. The use of either computed sum scores or estimated factor scores as a general practice is continuously debated [@mcneish2020; @widaman2022; @mcneish2023; @widaman2024]. The use of either appears to largely depend on the researcher's particular research objective, theoretical suppositions, and necessity. Another common concern is the inappropriate treatment of ordinal variables as interval which levy strong assumptions, e.g., each item contributes equally to the variable being measured and that ordinal response categories are equidistant as though on an interval scale [@wang2013; for contrast, see @robitzsch2020].

A set of five survey items measured public trust in elections by inquiring into the degree of confidence that votes will be counted as voters intend, confidence that electoral systems are secure from technological threats, perceived commitment of election staff, confidence that outcomes will be fair, as well as confidence that the voting process will be fair. The items measuring trust and confidence in elections include:

-   How confident are you that votes in Maricopa County, AZ will be counted as voters intend in the elections this November?[^8]
-   How confident are you that election officials, their staff, and volunteers in Maricopa County, AZ will do a good job conducting the elections this November?
-   Think about the election staff and volunteers who handle the administration and conduct of elections in Maricopa County, AZ. How committed do you think they will be to making sure the elections held this November are fair and accurate?
-   How confident are you that the voting process will be fair in Maricopa County, AZ?
-   How confident are you that election systems in Maricopa County, AZ will be secure from hacking and other technological threats?

[^8]: Note that confidence, or lack thereof, that votes will be counted as voters intend could easily be construed as an expectation of electoral fraud. It is a given that the votes will be counted, but the question is whether they will be counted *as intended*. Although fraud isn't provided as the reason for why votes wouldn't be counted as intended, it also isn't implied that miscounts would be due to some negligent mistake. A nefarious insinuation might be somewhat implicit, but perhaps only to those who are inclined to suspect as such. A lack of confidence here doesn't necessarily impart the reason for that lack of confidence, which makes it suitable as a survey item to measure the presence of trust in election administration rather than an expectation of fraud.

Each item in this series presented four response options from "Not at all [confident/committed]", "Not too [confident/committed]", "Somewhat [confident/committed]", and "Very [confident/committed]". In total, six survey items measure different aspects of trust and confidence in election administration. From these items, I generated a composite score for each respondent that represents their variable level of trust in elections.

### Expecatation of Electoral Fraud

A series of five survey items captured *expectation of electoral fraud*. These items were prefaced with the question, "How likely do you think any or all of the following will happen during this year´s elections in Maricopa County, AZ?" Each item presented four response options: "Not likely at all", "Not too likely", "Somewhat likely", and "Very likely".

-   There will be voter fraud, that is, people who are not eligible to vote will vote, or vote more than once
-   Many votes will not actually be counted
-   Many people will show up to vote and be told they are not eligible
-   A foreign country will tamper with the votes cast in this area to change the results
-   Election officials in Maricopa County, Arizona will try to discourage some people from voting

Similar to the items capturing trust in elections, I computed a composite score for each respondent reflecting their expectations of electoral fraud.

## Model and Data Specification

The outcome of interest concerns both public trust in election administration and the expectation of electoral fraud. Because these variables are assessed by multiple Likert items, conventional methods call for combining the items of each set into separate Likert scales, also known as a summated rating scales.
However, many of the items comprising the trust scale directly oppose items capturing expectations of electoral fraud. For instance, one item has the respondent report their confidence that votes will be counted as voter's intend; whereas a contrasting item has the the respondent report their expectation that many votes will not be counted. These two items directly oppose each other, but there's no guarantee that individual's responded as such.

Since many of the items are not symmetric---i.e., response options are not bipolar with contrasting positive and negative response options---then a "trust in elections" scale could only be understood as a measure of trust in the positive use of the term (i.e., unipolar). Likewise, the expectation of fraud items fit to a scale that corresponds to a unipolar measure of *confident distrust* so to speak.

I would expect composite sum or mean scores, and item responses, on these two measures to be inversely correlated among the sample, but not mutually exclusive. Indeed, this is what I find.

I first constructed two multi-item composite scales of both *trust in elections* and *expectation of fraud*. Cronbach's internal consistency reliability coefficient \alpha were estimated for both scales and were adequate (\alpha > 0.8); the trust in elections scale \alpha = 0.93 and the expectation of electoral fraud scale \alpha = 0.86. In other words, the items in each of the scales were intercorrelated well enough to compose a single scale of a (presumed) unidimensional construct. I then plotted separate histograms of scores across the sample to visualize the distribution for each seen in @fig-dist. Note that for both scales, higher values indicate less expectation of electoral fraud and greater trust in elections, respectively.

```{r}
#| label: fig-dist
#| fig-cap: "Histograms of Composite (mean) Scores for Trust in Elections And Expectation of Electoral Fraud"

# trust scores are not normally distributed, bi-modal, neg skew
trst.scores.hist <- ggplot(data, aes(trst.mean.scores))+
  geom_histogram(aes(y = after_stat(density)), bins = 15,
                 color = 1, fill = "grey")+
  geom_density(lwd = 1.5, color = 1, fill = 1, alpha = 0.35)+
  geom_segment(x = round(mean(data$trst.mean.scores), 2), y = 0, yend = 0.80, color = "red")+
  labs(
    x = "Trust in Elections Mean Scores",
    y = "Density",
    caption = "mean = 3.09, sd = 0.69, median = 3.0, skew = -0.67, kurtosis = 0.25"
  )+
  annotate("text", x = round(mean(data$trst.mean.scores), 2), y = 0.85,
           label = paste("mean =", round(mean(data$trst.mean.scores), 2)), color = "red")

# eef scores are not normally distributed, also bi-modal pos skew
eef.scores.hist <- ggplot(data, aes(eef.mean.scores))+
  geom_histogram(aes(y = after_stat(density)), bins = 15,
                 color = 1, fill = "grey")+
  geom_density(lwd = 1.5, color = 1, fill = 1, alpha = 0.35)+
  geom_segment(x = round(mean(data$eef.mean.scores), 2), y = 0, yend = 0.80, color = "red")+
  labs(
    x = "Exp. of Electoral Fraud Mean Scores",
    y = "Density",
    caption = "mean = 2.27, sd = 0.75, median = 2.2, skew = 0.19, kurtosis = -0.55"
  )+
  annotate("text", x = round(mean(data$eef.mean.scores), 2), y = 0.85,
           label = paste("mean =", round(mean(data$eef.mean.scores), 2)), color = "red")

# place plots side by side
gridExtra::grid.arrange(eef.scores.hist, trst.scores.hist, ncol = 2)
```

Given the non-normal (or at least un-equal skew of the) distribution of the two scale scores and ordinal nature of the variable items, I conducted a Spearman's rank correlation test [@spearman1907] and found a negative correlation of $\rho$ -0.49 (Kendall's \tau = -0.39) between summated scores on the two scales.

The negative correlation between the items and the composite scores makes sense. Theoretically speaking, a person who is certain that electoral fraud will occur clearly lacks trust in the election. That is, expectation of electoral fraud does imply lack of trust in elections. However, the reverse isn't also necessarily true; it is reasonable to hold a degree of trust in elections alongside expectations that electoral fraud will occur. Holding both considerations isn't necessarily an example of cognitive dissonance, as a lack of confidence in elections doesn't necessarily imply an expectation of electoral fraud. A lack of trust in elections might render one open to potentialities, but expectation of fraud is in no way necessitated by lacking trust in elections. When considered along a spectrum, both can be considered as degrees of *confidence*; and in the absence of confidence, insecure.

Accordingly, those whose responses to the "trust in elections" items would be cancelled out by their responses to the "expectation of fraud" items. Those who are, or are close to being, equally confident in both directions would be the most *insecure* about their expectations[^Insecure in terms of their confidence. Not insecure as to imply some anxious or concerned affect.]. Having no trust in elections and no expectation that electoral fraud will occur is functionally equivalent to having equal degree of both. Quite literally, an uncertain feeling of "it could go one way or the other". Operationally, this would reflect a middling 0 on a scale of *confidence*. Seeing as how expectation of electoral fraud is also a level of confidence, then it is reasonable to place all items onto a symmetric scale where positive values represent trust in elections and negative values represent expectation of electoral fraud.

The primary dependent variable is *confidence in elections*, which reflects one's trust in election administration minus their expectations of electoral fraud. Operationally, the dependent variable is a linear composite scale of item responses across items measuring trust in elections and expectations of electoral fraud re-scaled to range from -3 to 3. Items comprising trust in elections were positively coded to range from 0 "Not at all confident/committed" to 3 "Very confident/committed". Likewise, items capturing expectation of electoral fraud were coded negatively to range from 0 "Not likely at all" to -3 "Very likely".

Computed mean or sum scores for multiple Likert items are commonly acceptable, as they are relatively robust in spite of violating assumptions of a normal distribution of data (or residuals).
due to the somewhat non-normal distribution of the sum or mean scores and the ordinal nature of the variable items, an ordered regression analysis is preferred.

The primary independent variable is the prospect of military veterans as election workers (i.e., staff or volunteers).

@@

The key question is whether trust in elections (or voter confidence) is response to the treatment vignette.

Note that the question wording for the survey items derived from the 2022 SPAE are modified slightly but substantially. The question wording in this study is most similar to those in 2022 SPAE that assess confidence that votes will be counted as voters intend without specific reference to their own vote (e.g., in their county, state, or nationwide). In this study, two survey items ask about confidence in vote counts in one specific county in Arizona and another referencing their local community. Moreover, the 2022 SPAE includes "I don't know" as a response option, whereas the comparable item in this study did not include a "don't know" response option. Comparison between responses to the two similar items was conducted....

@@

The median age was 46 (mean age was 47). There was 51.7% (n = 658) women, 47% (n = 598) men, and approximately 1.3% who identified as either Non-binary/third gender (n = 7) or preferred not to say (n = 9). A large proportion of the sample identified as White or Caucasian (n = 975, 76.65%), while all other non-White respondents comprised 27.83% of the sample. Those who held a graduate level degree (e.g., Master's, Doctorate, or Professional level) comprised 13% of the sample; those with either degree at the Associate or Bachelor's level comprised 36.24%, while 22.25% had some college but no degree; and 28.46% had either a high school level or equivalent education or less than high school. The largest proportion of the sample identified as Democrat at 44.64%, followed by Republicans at 42.59%. The proportion of true Independents was 12.78%.

@@

[^2]: There's little that could justify a conceptual distinction between public trust pre-election and public trust post-election. The temporal element renders the difference between pre- and post-election more operative than conceptual at a foundational level.

[^3]: Note that *electoral anxiety* is used in place of electoral trust/confidence simply to emphasize the anticipatory nature of the attitudes in question. Election anxiety is not taken to be the inverse of trust and confidence. A lack of trust does not necessarily imply (a state of) anxiety.
